More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-02-01 2016-Weak-FH-Bisim/WeakBisim.tex
Moved Week-fh-bisim to github

\def\AlgT{\mathcal{T}}
\def\AlgE{\mathcal{E}}
\def\AlgA{\mathcal{A}}
\def\AlgAS{\mathcal{A}_S}
\def\AlgB{\mathcal{B}}
\def\AlgI{\mathcal{I}}
\newcommand{\set}[1]{\overline{#1}}
\newcommand{\Pred}{\symb{Pred}}
\newcommand{\Post}{\symb{Post}}
%\usepackage[math]{cellspace}
%\setlength\cellspacetoplimit{ 37pt}
%\setlength\cellspacebottomlimit{18pt}

@@

%\newtheorem{theorem}{Theorem}[section]
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{cor}[theorem]{Corollary}
%\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{algorithm}[theorem]{Algorithm}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{example}[theorem]{Example}
%\newtheorem{problem}[theorem]{Problem}
%\newtheorem{proof}[theorem]{Proof}

@@

\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\Flat}{Flat}
\renewcommand{\P}{\mathcal P}

\title{A Theory for the Composition of Concurrent Processes\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{ Ludovic Henrio\inst{1}  \ \ \  Eric Madelaine\inst{1,2} \ \ \ Min Zhang\inst{3}}
\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
\and Shanghai Key Laboratory of Trustworthy Computing, ECNU, China}
\date{}                                           % Activate to display a given date or no date

@@

\begin{abstract}
 In this paper, we provide a theory for the operators composing 
 concurrent processes.  
   Open pNets  (parameterised networks of synchronised automata) are
  new semantic objects that we propose for defining the semantics of
  composition operators. This paper defines the  operational semantics of open pNets, 
  using ``open
  transitions'' that include symbolic hypotheses on the behaviour of the
  pNets ``holes''. 
  We discuss when this semantics can be finite and how
  to compute it symbolically, and we illustrate this construction
  on a simple operator. This paper also defines a
   bisimulation equivalence between open pNets, and shows its decidability.
%  \TODO{chose between british and en\_US spelling (e.g. synchronised / synchronised)}

@@

 In the nineties, several 
works extended the basic behavioural models based on labelled
transition systems to address value-passing or parameterised systems, using
various symbolic encodings of the
transitions~\cite{deSimone85,Larsen87,HennessyLin:TCS95,Linconcur96}. 
In \cite{Linconcur96}, H.M. Lin addressed value-passing calculi, for which he
developed a symbolic behavioural semantics, and proved algebraic properties.
Separately J. Rathke~\cite{HennessyRathke:TCS98} defined another
symbolic semantics for 
a parameterised broadcast calculus, together with strong and weak bisimulation
equivalences, and developed a symbolic model-checker based on a tableau
method for these processes. 30 years later, no
practical verification approach and no verification platform are
using this kind of approaches to provide proof methods for
value-passing processes or open process expressions. 
This article proposes a new approach to   study concurrent and
distributed systems based on a semantic formalism featuring:
1) low-level description of behaviours (transition systems) with explicit
data parameters, and hierarchical structure, 2) flexible composition
and synchronisation mechanism, 3) finite representation of the
behavioural semantics using symbolic representations of sets of behaviours.

Parameterised Networks of synchronised automata (pNets) was proposed
to give a behavioural specification formalism for distributed systems.
It inherited from the work of
Arnold on synchronisation vectors~\cite{Arnold1982}. 
% The pNet model is  built on top of parameterised labelled
% transition systems(pLTS), extended with explicit data handling, and with
%hierarchical structuring mechanism.
In previous work~\cite{HMZ:PDP15}, we
showed that pNets can be used to represent the behavioural semantics
of a system 
including value-passing and many kinds of synchronisation methods. We
used these results to give the semantics of various constructs and
languages for distributed objects, and to build a platform for design
and verification of distributed software components
\cite{CM:FMCO08,HKM-FASE16}. 
The parameterised and hierarchical nature of pNets allows for compact
models easy to generate from applications in high-level languages.
Their structure is static, but unbounded, and this allows for model-checking
approaches even for reconfigurable applications.
Closed pNets were used to encode fully defined programs or systems,
while open pNets have ``holes'', playing the role of process
parameters. Such open systems can be used to define composition operators.
The challenge raised by the research on open pNets is due to its
``open'' nature and to the existence of holes and parameters.

\paragraph{Contribution}
The aim of this paper is to  provide a theory for the operators
composing concurrent processes. This theory is based on the definition of operators 
as open pNets. By defining the operational semantics of open pNets, using
open  transitions that include symbolic hypotheses on the behaviour of
the pNets holes, we  can define a strong bisimulation equivalence
between open pNets, and show its decidability. This work highlights the possibility to 
automatically infer proof obligations, in the form of predicate inclusion, that have to 
be verified to prove the equivalence of operators. These results allow us to envision the 
semi-automatic proof of equivalence between operators for composing processes.

@@

A number of fundamental works have been published on symbolic or open
bisimulations, with varying  vocabulary. In this section, we only list works that are
directly related to our approach.

The closest research (and oldest) is from R. De Simone
\cite{deSimone85}, who defines {\em Specification
Rules} and a {\em FH-bisimulation} equivalence, that were one of our main inspiration for 
the open-transition
concept.
Some years later, A. Rensink \cite{Rensink97} defines a generic
notion of conditional
  transition systems and studies relations between FH-bisimulation and
  others. We believe that in the work of De Simone context and in ours, the relations
  coincide, and that Rensink work differs mainly in presence of
  recursive binding constructs that we do not consider.

 In~\cite{HennessyLin:TCS95,Linconcur96} M. Hennessy and H.M. Lin developed
 the theory of symbolic transition graphs (STG), and the associated
 symbolic (early and late) bisimulations, they also study STGs with assignments which can 
 be a 
 model for message-passing processes. These are clearly related to our
 parameterised LTSs, though they are more specifically addressing the
 action algebra of value-passing
 CCS expressions. \cite{HennessyLin:TCS95} also gives an algorithm for
 computing symbolic bisimulation, but only for symbolic finite trees.
 %% \item Later,
 %% in~\cite{LinFORTE98},  H.M. Lin proposed a technology using 
 %% "instantiate input variables on-the-fly" which can deal with
 %% infinite value domain based on the STGs.
 An interesting variant was developped by Hennessy, and
 Rathke \cite{HennessyRathke:TCS98}, concerning a calculus of
 broadcasting systems (CBS) and a symbolic bisimulation. The main characteristic of 
 this calculus is that communication is
 ``one-to-many'', and non blocking, so the definitions of semantics
 and equivalences differ significantly from previous works.
 Later, J. Rathke proposed 
 a model-checker for CBS based on a sound tableau method over
 symbolic graphs.  
 Another important similarity between the works on STGs, CBS, and
 ours is the use of an auxiliary proof system on
 value expressions. Remark that pNets can encode both value-passing
 CCS and CBS, but also other communication and synchronisation schemas.

More recently, Y.X. Deng~\cite{Deng2001} gave an open bisimulation
 for $\pi$-calculus based on STG, which used a 
 predicate equation system whose greatest solution characterizes the
 condition under which the two STGs are bisimilar. There is here a
 potential relation with our work: if the number of states and
 branching  of the symbolic model is finite, then their algorithm can
 terminate; a similar approach may help us to compute our 
 FH-bisimulation.

    %% For the value-passing  processes, one solution is to give a
    %% symbolic transition graph model.   In the ~\cite{Linconcur96},
    %% H.M. Lin extended  the symbolic transition graph with assignments
    %% which can be a model for message-passing processors. Based on this
    %% model, he gave a symbolic operational semantics which is helpful
    %% for computing the symbolic bisimulation. In the
    %% ~\cite{LinFORTE98}, H.M. Lin proposed the technology  of
    %% "instantiate input  variables on-the-fly" which can deal with
    %% infinite value domain based on the STGs. In the ~\cite{Deng2001},
    %% Y.X. Deng gave another open bisimulation based on STG, which is
    %% different from our open automata. They used the predicate equation
    %% system whose greatest solution characterizes the condition under
    %% which  the two STGs are bisimular.

 %% \item\TODO{to be checked...}
%% \TODO{Min ??? missing cav97 paper... DONE}
%%   T.Bultan etc~\cite{Bultan97} proposed a model checker to verify the safety
%% and liveness properties on infinite-state programmes. The idea they used is
%% to symbolically encodes  transition relation and sets of states using affine
%% constraints  on integer variables, logical connectives and quantifiers and
%% formed the state-partitions over Cartesian-product of all variable domains.

Finally, there are numerous works on subclasses of
infinite-state programs or parameterised systems, seeking
decidability properties, and sometimes model-checking or equivalence
checking algorithms.
For example~\cite{Bultan97} proposed a model
checker to verify the safety and liveness properties on infinite-state
programs. They symbolically encode transitions and states using
predicates, including affine constraints on integer variables.
Another very different approach is used by 
\cite{ClarkeGJ97}, relying on a dedicated model based on network grammars
and regular languages.

\paragraph{Structure.}
Section \ref{section:pnets} extends the previous definition of pNets~\cite{HMZ:PDP15} to 
fit the needs of the open
pNets. Section~\ref{section:op-semantics} gives their operational
semantics based on open transitions, and proves that this semantics is finite under 
reasonable
conditions. In Section \ref{section:bisimulation} we introduce an
equivalence called FH-bisimulation, and prove its decidability.
All sections are illustrated by a running example encoding a Lotos operator.
Finally Section \ref{section:conclusion} concludes and discusses future work.

@@

@@

\paragraph{Alternative syntax.}
When describing examples, we usually deal with pNets with finitely
many sub-pNets and holes, and it is convenient to have a more concrete
syntax for synchronisation vectors. When $I\cup J\!=\!\![0..n]$  we
denote
% the pNet as $\mylangle \pNet_{1}, \ldots,
% \pNet_n,\symb{SV}\myrangle$,  and
synchronisation vectors as
$<\alpha_1,..,\alpha_n> \to\!\alpha$, and elements not
taking part in the
synchronisation  are denoted $-$ as in: $< -, -, \alpha, -, - > \to\!
\alpha$.

@@

\begin{example}

	To give simple intuitions of the open pNet model and its semantics, we use
	here a small  example coming from the Lotos specification language. It will be used 
	as  an illustrative example in the whole paper.
	We already have shown in \cite{HMZ:PDP15} how to encode non trivial
	operators using synchronisation vectors and one or several pLTSs used
	as controllers, managing the state changes of the operators.
	In Fig. \ref{schema:enable-pnets}, we show 2 possible encodings of the Lotos 
	``Enable''
	operator.
	In the Enable expression ``P>>Q'', an \texttt{exit(x)} statement within P
	terminates the current process, carrying a value $x$ that
	is captured by the \texttt{accept(x)} statement of Q.

	We use a simple action algebra,
	containing two constructors $\delta(x)$ and $acc(x)$, for any
	possible data type of the variable $x$, corresponding to the
	statements \texttt{exit(x)} and \texttt{accept(x)}. Both $\delta(x)$
	and $acc(x)$ actions are
	implicitly included in the sorts of all processes.
	We need no specific predicate over the action expressions, apart
	from equality of actions. 
	In the first encoding $Enable1$, in the upper part of
	Fig. \ref{schema:enable-pnets}, we use a controller $C_1$ with two
	states, and simple control actions $l, r, \delta$. The second encoding
	$Enable2$ uses a data-oriented style, with a single state controller, and a
	state-variable $s_0$, with values in $\{0,1\}$.

        In this example we use a specific notation for \emph{local
          actions}, that cannot be further synchronised, like the
        $\tau$ silent action of CCS. We name them \emph{synchronised
          actions}, and denote them as any action expression with the text
        underlined, as e.g. $\underline{\delta(x_2)}$. Such
        synchronised actions do not play any special role for defining
        strong bisimulation, but as one can expect, will be crucial
        for weak equivalences.

	Note that synchronisation vectors are defined in a parameterised
	manner:  the first and third lines represent one vector
	for each parameterised action in the Sort of hole $P$
	(resp. $Q$). This notation can also use predicates, as in the first
	case, in which we want the vector to apply to any action of $P$
	except $\delta(x)$.

	In Fig. \ref{schema:enable-composed}, we enrich our example by
	composing 2 occurences of the Enable1 pNet. To simplify,
	we only have represented one instance of the synchronisation vector set,
	and of the controller.
%	
%	We will use these pNets to
%	illustrate the formal definition in the next section, then their
%	operational semantics and their equivalence.

	The reader can easily infer from these two figures the
	following sets:

~~~~	{\small
$\begin{array}{l}
			\Holes(EnableCompL) = \{P,Q,R\}\\		
		\Leaves(EnableCompL)) = \{C_3,C_4\}\\	
		\Sort(C_1) = \Sort(C_2) = \Sort(C_4) = \{l,\delta,r\})\\		
		\Sort(EnableCompL) = \Sort(P)\backslash\{\delta(x)\}\cup
		\Sort(Q)\backslash\{\delta(x)\}\cup  \Sort(R)
		\cup \{\underline{\delta(x)}\} %,\underline{\delta(y)}\}_
		%{\delta(x)\in\Sort(P), \delta(y)\in\Sort(Q)}

          \end{array}$	}

%	We shall continue using these pNet examples to illustrate the
%	behavioural semantics construction, and the bisimulation equivalence.
\end{example}

@@

In \cite{HMZ:PDP15} we defined an operational semantics for closed
pNets, expressed in a {\sl late style}, where states and transition
were defined for a specific valuation of all the pNet variables.
Here we have a very different approach: we build a direct symbolic
operational semantics for open pNets,
encoding formally hypotheses about the
behaviour of the holes, and dealing symbolically with the variables.
This will naturally lead us in the following sections to the
definition of an open bisimulation equivalence, playing explicitly
with predicates on the action of holes, and values of variables.

The idea is
to consider an open pNet as an expression similar to an open process
expression in a process algebra. pNet expressions can be combined to
form bigger expressions, at the leaves pLTSs are constant expressions,
and holes play the role of process parameters.
In an open pNet, pLTSs naturally have states, and holes have no state;
furthermore, the shape of the pNet expression is not modified during
operational steps, only the state of its pLTSs can change.

@@

%%
%%
%%Each synchronisation vector $\symb{SV}_k$ has the form
%%$(\!\alpha_{m}^{m\in I_k}, \alpha_{j}^{j\in J_{k}})\to\alpha'$
%%with $I_k$ ranging over its sub-pNets, and $J_k$ ranging over its holes.
%%
%%The actions $\alpha_{m}$, $\alpha_{j}$, and $\alpha'$ are expressions from the action
%%algebras potentially containing variables (data parameters) relating
%%the actions of the involved sub-pNets and the resulting actions. These
%%relations are conditions for this synchronisation to
%%happen: the synchronisation occurs when a unification of the sub-pNets
%%actions with the corresponding vector actions exists; this unification
%%yields an instanciation, that is applied to the global action of the
%%vector $\alpha'$.
%%
%%\TODO{Something is missing here... we remove the formal def pf
%%  predicates, but we should have something short instead.}
%% \begin{definition}[Synchronisation Predicate]
%% Let $V_k$ be the set of variables in the action expressions of $\symb{SV}_k$.
%% We call {\sl synchronisation predicate}, and denote as
%% $Pred(V_k, a_i^{i\in I'}, b_j^{j\in J'}, v)$
%% the predicate relating the actions of the leaves and holes
%% involved in $\symb{SV}_k$ with its resulting (global) action. This predicate
%% will be used in the SOS rule defining the bahaviour of a pNet.
%% \end{definition}
%\begin{definition}[Open transitions]
%\label{def:OpenTransitions}
%For any open \pNet\  $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in
%  K}\myrangle$, let $s_i, s_i^{\prime}$ range over the states of its set of leaves
%$pLTS_i^{i\in L}$, and let $Holes_j^{j\in J}$ denote its set of holes.
%An \emph{open transition} of this open \pNet \ is a rule:
%   \begin{mathpar}
%     \inferrule*[myfraction=\reddottedrule]
%         {\{s_i\xrightarrow{a_i} s_i^{\prime}\}_{i\in L^\prime},
%           \{\xrightarrow{b_j}\}_{j\in J^\prime}, \Pred(SV_k, a_i^{i\in L'}, b_j^{j\in
%           J'}, v)}
%         {\ostate{s_i^{i\in L}} \xrightarrow {v} \ostate{s_i^{\prime i\in L}}}
%\end{mathpar}
%in which $L^\prime\subseteq L$ and $J^\prime\subseteq J$ are the sets
%of leaves and holes {\bf involved} in this rule, and for
%$i\notin L'$, $s_i^\prime = s_i$.  $\Pred(SV_k, a_i^{i\in L'}, b_j^{j\in J'}, v)$ is a
%predicate relating the actions of the leaves and holes
%involved in $\symb{SV}_k$ with its resulting (global) action.
%\end{definition}
%%% \TODO{question: is this rule GSOS format compliant ? I suppose no, because
%%%   of our complex sets/indexes syntax, and GSOS is about syntax... Different manner to
%%%set this question, considering 2 different point of views: the open-transition is a
%%%structure (symbolic) of the automaton model; while the specification rule is an
%%%artifact
%%%used for building derivation trees, that can be compared to GSOS style formats.}
%We build the semantics of open pNets as a set of open transitions,
%that we call an {\bf open automaton}. States are the states of the pNet
%expressions, so they only encode the states of its leaves. The pNet
%structure itself does not appear in the open-automaton structure, only a
%set of Holes and a set of Leaves.
%\begin{definition}[Operational semantics of open pNets]
%  \label{def:operationalSemantics}
%The semantics of a pNet $p$ is an open automaton $A = <L,J,\overline{\mathcal{S}},
%\mathcal{S}_0, \mathcal{T}>$ where:
%\begin{itemize}
%\item $L$ is the indices of the leaves of $p$: $Leaves(p) =
%  \pLTS_i^{i\in L}$; and $J$ the indices of the holes: $Holes(p)= H_j^{j\in J}$.
%%  \item $\overline{L}^L = Leaves(p), \overline{H}^J = Holes(p)$
%  \item $\overline{\mathcal{S}} = States(p)$
%  \item $\mathcal{S}_0 = InitState(p)$
%  \item $\mathcal{T}$ is the smallest set of open transitions
%      satisfying the rules below:
%\end{itemize}
%\TODO{First version, without handling of local variables and
%  assignments. We should be careful here: after (re) reading "Huimin
%  Lin, 'Symbolic Transition Systems with Assignements', Concur'96" I
%  think handling assignments is not trivial, even for comparisons of pLTSs. }
%The rule for pLTS (Tr1) is straightforward:
%\begin{description}
%  \item[{\bf Tr1:}]
%    $\inferrule
%    { s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
%      J}\rangle} s'\in \to }
%   { p = \mylangle  S,s_0, \to \myrangle
%     \models
%     \inferrule*[myfraction=\reddottedrule]
%         {\{s \xrightarrow{\alpha} s'\} \quad \emptyset \quad e_b}
%         {\ostate{s} \xrightarrow{\alpha} \ostate{s'}}
%   }
%               $
%  \end{description}
%The second rule deals with pNet nodes: for each possible
%synchronisation vector applicable to the rule subject, the premisses
%include one {\em open transition} for each sub-pNet involved, one possible
%{\em action} for each Hole involved, and the predicate relating these
%with the resulting action of the vector.
%A key to understand this rule is that the open transitions are
%expressed in terms of the leaves and holes of the pNet structure,
%i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
%Leaves, $L_k$ the index set of the leaves of one subnet, so all $L_k$
%are disjoint subsets of $L$. Thus the states in the open transitions,
%at each level, are tuples including states of all the
%leaves of the pNet, not only those involved in the chosen
%synchronisation vector.
%   \begin{description}
%   \item[{\bf Tr2:}]
%   \end{description}
%    \medskip\noindent
%    in which :\\
%- all variables in the synchronisation vector $SV_k$ are fresh\\
%- the resulting $Pred$ is a predicate collecting the constraints from the premisses of
%the rule, namely:
%    \noindent
%    $Pred = \bigwedge_{m\in I_k} Pred_m \land \bigwedge_{m\in I_k} v_m = \alpha_m \land
%\bigwedge_{j\in J_k} (b_j = \alpha_j \land b_j \in Sort_j) \land v = \alpha_k'$
%    \medskip
%%%    \TODO{I have tentatively added the sort constraint on hole actions, that was not
%%%included in the first version... I'm unsure whether this is the best place to include
%%%it,
%%%because it may change the decidability conditions on predicates}
%\end{definition}

@@

@@

2a) pick one state from the unexplored set and add it to the explored
set. From this state
build all possible deduction trees by application of the structural
rules Tr1 and Tr2, using all applicable combinations
of synchronisation vectors.

@@

2c) Optionally, simplifying the predicate at this point may minimize the
resulting transitions, or even prune the search-space.

2d) For each open-transition from step 2b, add the
resulting state in the unexplored set if it is not already in the
explored set, and add the transition in the outgoing transitions of
the current state.

@@

\begin{proof}
  The possible set of states of the open-automaton is the cartesian
  product of the states of its leaves $pLTS_i^{i\in L}$, that is finite by
  hypothesis. So the top-level residual loop of Algorithm 1 terminates provided each 
  iteration terminates. The enumeration of open-transitions in step 2b is bounded by the
  number of applications of rules Tr2 on the structure of the
  pNet tree, with a finite number of synchronisation vectors applying
  at each node the number of global open transition is finite. Similarily rule Tr1 is 
  applied finitely if the number
  of transitions of each pLTS is finite. So we get finitely many deduction trees, and
  open-transitions which ensures that each internal loop of Algorithm 1 terminates.
  \qed
  \end{proof}
%  
%\begin{theorem}[ Finiteness of open-automata.]
%
%Given an open pNet $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in 
%K}\myrangle$ with leaves $pLTS_i^{i\in L}$ and holes $Hole_j^{j\in
%  J}$, if $ finite{(L)}$ and  $ finite{(J)} $ and $ finite{(K)} $  and 
%$\forall i \in L. $,  $ finite{(states(pLTS_i))} $ and $ pLTS_i$
%has a finite number of state variables, then Algorithm 1 terminates
%and produces an open automaton 
%  $\mathcal{T}$ with finitely many states and transitions.
%
%
%
%\end{theorem}
%
%\begin{proof}
% Min's :  The set of states of the open-automaton is the cartesian product of the states 
%of its leaves $pLTS_i^{i\in L}$ . It is finite because of hypothesis. According to the 
%algorithm 1, starting with the initial state of the automaton, it can be terminate to 
%build all possible deduction trees by rules Tr1 and Tr2.  Because each pLTS is finite, 
%the number of applications of Tr1 is finite; because the set of synchronisation vectors 
%is finite ($ finite{(K)} $ ) , the number of applications Tr2 is finite. So we can get 
%finite open transitions. 
%  \end{proof}

%% \subsection{Simplication of predicates}
%% \TODO{Two facets here: \\
%% 1) The operational rules generate and assemble a large number of
%% predicate fragments, with many variables. It can be possible using a
%% solver to prove satisfiability, unsatisfiability, or to simplify these
%% predicates. This has 2 kinds of benefits: eliminate uneccessary
%% variables and reduce the predicate size; and prune the research space
%% when predicates are unsatisfiable.\\
%% 2) We have quite a specific form of building our predicates, with
%%   many simple equations having a fresh variable on the left. Only the
%%   predicate from the Sort membership is different... Can we
%%   derive a notion of normal form, or some simplification principle ?}.\\
%% Deciding what strategy using simlificatiom would be the most efficient
%% depends on the generation algorithm, and we need time for this.
%% For this paper, we can simply comment on these remarks, and leave
%% deeper study for future work.

@@

	Now we use our symbolic operational semantics to define a notion of
	strong (symbolic) bisimulation.Moreover this equivalence is decidable whenever we 
	have some
	decision procedure on the predicates of the action algebra.

@@

 \begin{definition}[FH-bisimulation]\label{def-FH-bisim}

\noindent
\begin{minipage}{0.69\linewidth} 	Suppose that
   $A_1 = <L_1,J,\mathcal{S}_1, s^1_0,
   \mathcal{T}_1>$ and $A_2 = <L_2,J,\mathcal{S}_2,s^2_0, \mathcal{T}_2>$
   are open automata where the set
   of holes are equal %\LUDO{added:} 
   and of the same sort.  %\LUDO{changed:} 
      Let 
      $(s^1,s^2|\Pred)\in\mathcal{R}$ be a
      relation over the sets 
      $\mathcal{S}_1$ and $\mathcal{S}_2$ 
      constrained by a predicate.  More precisely, for any pair $(s^1,s^2)$, there is a 
      single
      $(s^1,s^2|\Pred)\in\mathcal{R}$  stating that $s^1$ and $s^2$ are related if \Pred\ 
      is 
      true.
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
\end{minipage}
 Then 
   $\mathcal{R}$ is an FH-bisimulation iff for any  states
   $s^1\in\mathcal{S}_1$ and\\ $s^2\in\mathcal{S}_2$, $(s^1,s^2|\Pred)\in\mathcal{R}$, we 
   have
   the following:

 %  \TODO{Eric: Still have to check fresh variables in predicates}

@@

@@

\begin{theorem}[Context equivalence]\label{thm-ctxt-eq}
	Consider two FH-bisimilar open pNets:
	$\pNet = \mylangle \pNet_i^{i\in I}, S_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$ and 	$\pNet' = \mylangle {\pNet'}_i^{i\in I}, 
	S_j^{j\in 
	J}, 	\overline{\symb{SV'}}\myrangle$ 
	(recall they must have the same holes to be bisimilar).
	Let $j_0\in J$ be a hole, and $Q$ be a pNet such that $\Sort(Q)=S_{j_0}$. Then $\pNet[Q]_{j_0}$ and 
	$\pNet'[Q]_{j_0}$ are FH-bisimilar.
\end{theorem}
%
%In the following we use $\approx$ to denote equality modulo alpha conversion of free 
%variables. In practice, $x\approx y$ can be obtained if: $x=clone(a) \land y=clone(a)$.

The proof of theorem \ref{thm-ctxt-eq} relies on two main lemmas,
dealing respectively with the decomposition of a composed behaviour
between the context and the internal pNet, and with their recomposition. 
We start with decomposition: from one open transition of $P[Q]_{j_0}$, we exhibit 
corresponding behaviours of $P$ and $Q$, and determine the relation between their 
predicates:

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:

>>>>

\def\AlgT{\mathcal{T}}
\def\AlgE{\mathcal{E}}
\def\AlgA{\mathcal{A}}
\def\AlgAS{\mathcal{A}_S}
\def\AlgB{\mathcal{B}}
\def\AlgI{\mathcal{I}}

\newcommand{\Pred}{\symb{Pred}}
\newcommand{\Post}{\symb{Post}}
%\usepackage[math]{cellspace}
%\setlength\cellspacetoplimit{ 37pt}
%\setlength\cellspacebottomlimit{18pt}

@@

%\newtheorem{theorem}{Theorem}[section]
\newtheorem{prop}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}
%\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{algorithm}[theorem]{Algorithm}
%\newtheorem{remark}[theorem]{Remark}
%\newtheorem{definition}[theorem]{Definition}
%\newtheorem{example}[theorem]{Example}
%\newtheorem{problem}[theorem]{Problem}
%\newtheorem{proof}[theorem]{Proof}

@@

\title{More equivalences for Open pNets\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{ Ludovic Henrio\inst{1}  \ \ \  Eric Madelaine\inst{1,2} \ \ \ Min Zhang\inst{3}}
\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
\and Shanghai Key Laboratory of Trustworthy Computing, ECNU, China}
\date{}                                           % Activate to display a given date or no date

@@

\begin{abstract}
  Strong bisimulation allows to compare open systems described using the pNets model.
  In practice, as happens in process algebras, it is too strong, and
  we need to define some coarser relations, taking into account
  invisible or internal moves.

  Open pNets have a notion of \emph{synchronised actions}
  generalizing the usual internal actions (e.g. $\tau$ of CCS, or $i$
  in Lotos).

  We use this notion to define an equivalence relation similar to
  the classical \emph{weak bisimulation}, and study its properties...

  refinement preorders ?

  Algorithms ?

  Which significant (small) use-case ?

@@

\ERIC{imported from FORTE'16, but many of these works also had weak
  version... we need to have another look}

In the nineties, several 
works extended the basic behavioural models based on labelled
transition systems to address value-passing or parameterised systems, using
various symbolic encodings of the
transitions~\cite{deSimone85,Larsen87,HennessyLin:TCS95,Linconcur96}. 
In \cite{Linconcur96}, H.M. Lin addressed value-passing calculi, for which he
developed a symbolic behavioural semantics, and proved algebraic properties.
Separately J. Rathke~\cite{HennessyRathke:TCS98} defined another
symbolic semantics for 
a parameterised broadcast calculus, together with strong and weak bisimulation
equivalences, and developed a symbolic model-checker based on a tableau
method for these processes. 30 years later, no
practical verification approach and no verification platform are
using this kind of approaches to provide proof methods for
value-passing processes or open process expressions.

\paragraph{Contribution}

@@

\paragraph{Structure.}

@@

\ERIC{Keep minimum defs here too be independent from FORTE}

@@

@@

@@

@@

\end{definition}
%

@@

\TODO{Eric: refine the algorithm description}

@@

2a) pick one state from the unexplored set and add it to the explored
set. From this state
build all possible deduction trees by application of the structural
rules Tr1 and Tr2, using all \ERIC{applicable} combinations
of synchronisation vectors.

@@

2c)
\ERIC{The predicate is submitted to Z3 for checking satisfiability. If it is NOT stisfiable, the resulting OT is discarded. This will minimize the number of
resulting transitions, and potentialy prune the search-space.}

2d)
\ERIC{Optionally, simplify the predicate by eliminating the unnecessary intermediate variables, that were produced as fresh local variables of synchronisation vectors, and fresh result variables of intermediate OTs. In the resulting predicate the only significant variables are :\\
- the input variables of pLTS transitions\\
- the actions of holes\\
- the result action at toplevel.}

For each open-transition,  add the transition in the outgoing transitions of
the current state, and add the
resulting state in the unexplored set if it is not already in the
explored set.

@@

\section{Tooling}

\TODO{Eric: high level description of the implementation}

\ERIC{
=> Context VCE, ref to FASE paper\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule Tr2, the resulting OT is built from the premisses, without explicitely constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone'' operations, using structured variable names ensuring uniqueness of fresh names, but also reasonable readability of these names for debugging purposes.\\
=> Satisfiability check is done only at toplevel of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentialy reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
=> Simplification is not yet implemented. It is not strictly required for the Open Automaton construction, but it will be critical later for predicate comparison in the bisimulation algorithm.}

\TODO{Eric: full result of the tool execution on the running example}

@@

	Now we use our symbolic operational semantics to define a notion of
	strong (symbolic) bisimulation. Moreover this equivalence is decidable whenever we 
	have some
	decision procedure on the predicates of the action algebra.

@@

 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim}

\noindent
\begin{minipage}{0.69\linewidth} 	Suppose that
   $A_1 = <L_1,J,\mathcal{S}_1, s^1_0,
   \mathcal{T}_1>$ and $A_2 = <L_2,J,\mathcal{S}_2,s^2_0, \mathcal{T}_2>$
   are open automata where the set
   of holes are equal %\LUDO{added:} 
   and of the same sort.  %\LUDO{changed:} 
      Let 
      \ERIC{$\mathcal{R}=\{(s^1,s^2|\Pred_{s^1,s^2})\}$} be a
      relation over the sets 
      $\mathcal{S}_1$ and $\mathcal{S}_2$ 
      constrained by a predicate.  More precisely, for any pair $(s^1,s^2)$, there is a 
      single
      $(s^1,s^2|\Pred_{s^1,s^2})\in\mathcal{R}$  stating that $s^1$ and $s^2$ are related if $\Pred_{s^1,s^2}$\ 
      is 
      true.
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
\end{minipage}
 Then 
   $\mathcal{R}$ is an FH-bisimulation iff for any  states
   $s^1\in\mathcal{S}_1$ and\\ $s^2\in\mathcal{S}_2$, $(s^1,s^2|\Pred)\in\mathcal{R}$, we 
   have
   the following:

@@

\begin{theorem}[FH-Bisimulation is an equivalence]\label{thm-equiv} Suppose $\mathcal{R}$ 
is an FH-bisimualtion. Then $\mathcal{R}$ is an equivalence, that is, $\mathcal{R}$ is 
reflexive, symmetric and transitive.

\end{theorem}

@@

\LUDO{added congruence theorem below}
\begin{theorem}[Congruence]\label{thm-congr-eq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, S_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $Q$ and $Q'$ be two FH-bisimilar pNets such that 
	$\Sort(Q)=\Sort(Q')=S_{j_0}$\footnote{Note that $\Sort(Q)=\Sort(Q')$ is ensured by 
	stronfg bisimilarity.}. Then 
	$\pNet[Q]_{j_0}$ and 
	$\pNet[Q']_{j_0}$ are FH-bisimilar.
\end{theorem}

\begin{theorem}[Context equivalence]\label{thm-ctxt-eq}
	Consider two FH-bisimilar open pNets:
	$\pNet = \mylangle \pNet_i^{i\in I}, S_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$ and 	$\pNet' = \mylangle {\pNet'}_i^{i\in I}, 
	S_j^{j\in 
	J}, 	\overline{\symb{SV'}}\myrangle$ 
	(recall they must have the same holes to be bisimilar).
	Let $j_0\in J$ be a hole, and $Q$ be a pNet such that $\Sort(Q)=S_{j_0}$. Then $\pNet[Q]_{j_0}$ and 
	$\pNet'[Q]_{j_0}$ are FH-bisimilar.
\end{theorem}
